Nondeterministic Bigraphical Reactive Systems for Markov Decision Processes\thanks{Supported by organization x.}

### 2018-08-12 paper.tex
Adding some citations

%TODO: MDP generation: show, define

@@

\begin{definition}[concrete link graph]
  A \emph{concrete link graph}
  \[ F = (V_F, E_F, \ctrl_F, \link_F) : X \to Y \]
  is a quadruple having an inner face $X$ and an outer face $Y$, both finite
  subsets of $\mathcal{X}$, called respectively the \emph{inner} and
  \emph{outer names} of the link graph. $F$ has finite sets $V_F \subset
  \mathcal{V}$ of nodes and $E_F \subset \mathcal{E}$ of edges, a control map
  $\ctrl_F : V_F \to \mathcal{K}$ and a \emph{link map}
  \[ \link_F : X \uplus P_F \to E_F \uplus Y, \]
  where $P_F \coloneqq \{ (v, i) \mid i \in \ar(\ctrl_F(v)) \}$ is the set of
  ports of $F$. Thus $(v, i)$ is the $i$th port of node $v$. We shall call $X
  \uplus P_F$ the \emph{points} of $F$, and $E_F \uplus Y$ its \emph{links}.
\end{definition}

@@

TODO: Akin to \cite{dblp:conf/nfm/giaquintahimn18} and maybe
\cite{koenig1998xavier}.

@@

Our first scenario is concerned with collecting objects in a grid-like
two-dimensional world. The agent starts in the southwest corner of the map
(which we call \emph{home}), explores the environment until it collects a
specified number of objects, and returns home (while possibly collecting more
objects along the way). The model tracks which cells of the grid have been
explored, and only the unexplored cells have a constant probability $p$ of
rewarding the agent with an object. The desired number of objects can be easily
changed, and the map can be customised to make some transitions unavailable
(i.e. add walls), as long as the agent can still return home.

@@

TODO: Should mention \cite{DBLP:conf/giscience/WaltonW12}.

Our second model explores two important concepts in modelling the topology of
space: nesting and uncertainty. Nesting is the idea that spaces can be inside
other spaces: rooms inside floors, floors inside buildings, buildings inside
streets, etc. A key characteristic of nested spaces is that there is a separate
action (or several) taking you from the outer space to the inner space or vice
versa (e.g., using a door to enter or leave a room). In the rest of the paper,
whenever we discuss one space inside another, we will refer to the inner space
as the \emph{room}.

@@

As the ideas of collecting objects and tracking visited cells have been
explored already, they are dropped from this model to simplify the
implementation. The goal can simply be defined as moving from one location to
another. We add one extra control, \texttt{Node}. While \texttt{Cell} can be
imagined as a discrete patch of space (e.g., \SI{1}{\square\metre}),
\texttt{Node} can be thought of as a room. While we stick to two layers
(corridors and rooms) in this example, rooms can be freely nested: a room can
contain any number of rooms, each of which can contain more rooms.

@@

In order to implement nesting, a \texttt{Cell} contains a \texttt{Node}, which
can have any number of \texttt{Cell}s of its own. Since after entering a room,
you end up in a specific place inside the room (next to the door), we also need
to link up the \texttt{Node} with one of the \texttt{Cell}s it contains (see
Fig.~\ref{agent2_nesting}). Both \texttt{Cell}s and \texttt{Node}s support
having a single link for exactly this purpose. Thus, most of the \texttt{Cell}
ports are not linked to anything, which is implemented via closures (e.g.
\texttt{/x Cell\{x\}}). The initial map in Listing~\ref{agent2_initial} has a
corridor of three cells, with the agent on the west-most cell, and a door to a
room in the middle cell. One moves in and out of a room via a pair of actions
with reaction rules pictured in Fig.~\ref{go_in_out}.

@@

The room is a $2 \times 2$ grid of \texttt{Cell}s, which is generated when the
\texttt{Agent} gets to a \texttt{Cell} with an empty \texttt{Node} (represented
by \texttt{Node\{n\}.1})\footnote{Multiple different rooms can be implemented by
having each previously empty \texttt{Node} contain a node of different control.
This way a different probability distribution (with different possible outcomes)
can be assigned to each room.}. In order for this to happen before any other
action can take place, the two reaction rules in Listing~\ref{agent2_room} are
put on a higher priority class. With probability $p$, the generated room has
each \texttt{Cell} linked to its two neighbours; and with probability $1-p$, the
top two \texttt{Cell}s are not linked, possibly forcing the agent to choose a
different path. Finally, we have four actions, each with a single reaction rule
with probability $1$, corresponding to the four possible directions of travel.
The reaction rules are very similar to those of our previous example, see
Listing~\ref{agent2_north}.

@@

>>>>

@@

\begin{definition}[concrete link graph]
  A \emph{concrete link graph}
  \[ F = (V_F, E_F, \ctrl_F, \link_F) : X \to Y \]
  is a quadruple having an inner face $X$ and an outer face $Y$, both finite
  subsets of $\mathcal{X}$, called respectively the \emph{inner} and
  \emph{outer names} of the link graph. $F$ has finite sets $V_F \subset
  \mathcal{V}$ of nodes and $E_F \subset \mathcal{E}$ of edges, a control map
  $\ctrl_F : V_F \to \mathcal{K}$ and a \emph{link map}
  \[ \link_F : X \uplus P_F \to E_F \uplus Y, \]
  where $P_F \coloneqq \{ (v, i) \mid i \in \ar(\ctrl_F(v)) \}$ is the set of
  ports of $F$. Thus, $(v, i)$ is the $i$th port of node $v$. We shall call $X
  \uplus P_F$ the \emph{points} of $F$, and $E_F \uplus Y$ its \emph{links}.
\end{definition}

@@

In order to demonstrate the use of bigraphs in modelling probabilistic
processes with nondeterminism, we present two examples of agent movement
planning in two dimensions. The examples have different sources of probabilistic
behaviour and explore topics such as tracking visited locations and uncertainty
about the topology.

% TODO advantages: more visual/realistic than a vector of numbers, rewards and
% transition rules can be generated with a higher degree of automation, less
% code, more flexible (changing size of the grid); disadvantages: constant
% probability (without hacking)

% TODO xavier \cite{koenig1998xavier}

@@

Our first scenario is concerned with collecting objects in a grid-like
two-dimensional world and is inspired by similar scenarios modelled using PRISM
\cite{dblp:conf/nfm/giaquintahimn18,DBLP:conf/spin/HoffmannIMNV16}. The agent
starts in the southwest corner of the map (which we call \emph{home}), explores
the environment until it collects a specified number of objects, and returns
home (while possibly collecting more objects along the way). The model tracks
which cells of the grid have been explored, and only the unexplored cells have a
constant probability $p$ of rewarding the agent with an object. The desired
number of objects can be easily changed, and the map can be customised to make
some transitions unavailable (i.e. add walls), as long as the agent can still
return home.

@@

Our second model explores two important concepts in modelling the topology of
space: nesting and uncertainty. Nesting is the idea that spaces can be inside
other spaces: rooms inside floors, floors inside buildings, buildings inside
streets, etc. A key characteristic of nested spaces is that there is a separate
action (or several) taking you from the outer space to the inner space or vice
versa (e.g., using a door to enter or leave a room). In the rest of the paper,
whenever we discuss one space nested inside another, we will refer to the inner
space as the \emph{room}. While nesting has been modelled using bigraphs before
\cite{DBLP:conf/giscience/WaltonW12}, our approach models each room not as a
single entity, but rather as an interconnected collection of smaller spaces,
which we continue to call \emph{cells}. Thus, it is not enough to be in the
hallway in order to enter a room, one needs to be next to the door. Similarly,
after entering a room, the agent is positioned next to the other side of the
door.

@@

As the ideas of collecting objects and tracking visited cells have been
explored already, they are dropped from this model in order to simplify the
implementation. The goal can simply be defined as moving from one location to
another. We add one extra control, \texttt{Node}. While \texttt{Cell} can be
imagined as a discrete patch of space (e.g., \SI{1}{\square\metre}),
\texttt{Node} can be thought of as a room. While we stick to two layers
(corridors and rooms) in this example, rooms can be freely nested: a room can
contain any number of rooms, each of which can contain more rooms.

@@

In order to implement nesting, we put a \texttt{Node} (representing a room)
inside a \texttt{Cell}, which has the ``door'' to that room (see Fig.
\ref{agent2_nesting}). The \texttt{Node} can have its own collection of
\texttt{Cell}s, describing the topology of the room. One of these \texttt{Cell}s
is linked to the \texttt{Node}, representing the \texttt{Cell} closest to the
door from the inside of the room. Thus, both \texttt{Node}s and \texttt{Cell}s
have arity $1$ for exactly this linking. \texttt{Cell} ports not linked to
anything are implemented using closures (e.g. \texttt{/x Cell\{x\}}). The
initial map in Listing~\ref{agent2_initial} has a corridor of three cells with
the agent on the west-most cell and a door to a room in the middle cell. One
moves in and out of a room using a pair of actions with reaction rules pictured
in Fig.~\ref{go_in_out}.

@@

The room is a $2 \times 2$ grid of \texttt{Cell}s, which is generated when the
\texttt{Agent} gets to a \texttt{Cell} with an empty \texttt{Node} (represented
by \texttt{Node\{n\}.1})\footnote{Multiple different rooms can be implemented by
having each previously empty \texttt{Node} contain a node of different control.
This way a different probability distribution (with different possible outcomes)
can be assigned to each room.}. In order for the room to be generated before any
other action can take place, the two reaction rules in Listing~\ref{agent2_room}
are put on a higher priority class. With probability $p$, the generated room has
each \texttt{Cell} linked to its two neighbours; and with probability $1-p$, the
top two \texttt{Cell}s are not linked, possibly forcing the agent to choose a
different path. Finally, we have four actions, each with a single reaction rule
with probability $1$, corresponding to the four possible directions of travel.
The reaction rules are very similar to those of our previous example, an example
of which can be found in Listing~\ref{agent2_north}.

@@

BIG TODO!
